Binary Latent Representations for Efficient Ranking

### 2017-06-10 paper/paper.tex
Flesh out introduction.

%\usepackage{tabularx}

@@

\title{Binary Latent Representations for Efficient Ranking}
\author{Maciej Kula}
\date{\today}

\begin{abstract}
Large-scale recommender systems often face severe latency and storage constraints. To relax those, we train recommendation models that use binary rather than real-valued user and item representations, and show that they enable a range of attractive speed-accuracy trade-offs. In particular, we demonstrate 13 times faster prediction at the expense of a modest 4\% loss in ranking accuracy on the Movielens 100K dataset.
\end{abstract}

@@

\section{Introduction}

Industry ranking and recommendation systems need to scale to tens or hundreds of millions of items and users. Ranking latency (in on-line settings) and compute and storage costs (in pre-compute settings) place significant contraints on the design of the resulting system.

To alleviate these constraints, we propose the use of binary item and user representations in learning-to-rank matrix factorization models. Following the approach of \citet{rastegari2016xnor}, we show that we can estimate binary user and item representations that achieve comparable accuracy to standard learning-to-rank approaches while being several times faster to score and requiring a fraction of the memory. In our MovieLens 100k experiments, our binary representations are within 1\% of the accuracy of the real-valued model, but are over 2 times faster to score and take 50\% of the memory.

\section{Related work}
\citet{shrivastava2014asymmetric} propose a solution fo efficient ranking of large result sets through approximate maximum inner product search (MIPS). They extend well-known \citep{indyk1998approximate} locality-sensitive hashing techniques to the MIPS setting through asymmtric transformations of the query and candidate vectors, and show their applicability to recommender systems.

>>>>

@@

\title{Binary Latent Representations for Efficient Ranking}
\author{Maciej Kula}
\email{maciej.kula@gmail.com}
\date{\today}
\settopmatter{printacmref=false, printccs=false, printfolios=true}
\setcopyright{none}
\acmConference[]{}{}{}

\begin{abstract}
Large-scale recommender systems often face severe latency and storage constraints at prediction time. To relax those, we train recommendation models that use binary rather than real-valued user and item representations, and show that they enable a range of attractive speed-accuracy trade-offs. In particular, we demonstrate 13 times faster prediction at the expense of a modest 4\% loss in ranking accuracy on the Movielens 100K dataset.
\end{abstract}

@@

\section{Introduction}
Industry ranking and recommendation systems need to scale to tens or hundreds of millions of items and users. Ranking latency (in on-line settings) and compute and storage costs (in pre-compute settings) place significant contraints on the design of the resulting system.

In an on-line setting, where recommendation latency directly impacts the user experience, system designers have to work within a strict latency budget. Respecting that budget often means trading off model accuracy for speed, either via smaller and less expressive models (for example, with lower-dimensionality representations) or through aggressive use of heuristics to exclude large classes of candidate items from scoring.

The challenge of low-latency scoring leads many systems to rely on a pre-compute system, where recommendations are computed in advance and stored for serving at a later time. The primary disadvantage of such a system is its inability to react dynamically to user actions (such as new interactions that allow the recommendations to be refined) and recommendation context (location, time of day or year). Additionaly, while a pre-compute solution removes model prediction from the critical path, it still requires substantial investment in compute and storage resources to calculate and cache the predictions of the model.

To alleviate these constraints, we propose the use of binary item and user representations in learning-to-rank matrix factorization models. Following the approach of \citet{rastegari2016xnor}, we show that we can estimate binary user and item representations that achieve comparable ranking accuracy to standard learning-to-rank approaches while being several times faster to score and requiring a fraction of the memory. In MovieLens 100k experiments, our binary representations are within 4\% of the accuracy of the real-valued model, but are over 13 times faster to score and take 55\% of the memory.

\section{Related work}
\citet{shrivastava2014asymmetric} propose a solution for efficient ranking of large result sets through approximate maximum inner product search (MIPS). They extend well-known \citep{indyk1998approximate} locality-sensitive hashing techniques to the MIPS setting through asymmtric transformations of the query and candidate vectors, and show their applicability to recommender systems.
